---
title: "Tidy Tuesday Jan 14"
author: "Alice Tivarovsky"
date: "2025-01-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Introduction

This dataset explores posit::conf talks in 2023 and 2024. Some questions to think about: 
- Which speakers gave talks both years? 
- Are there keywords that appear in track titles in both 2023 and 2024? 
- What is the average sentiment of descriptions in each track? 

Be careful to de-duplicate the data when needed. Talks are indexed by speaker and some talks have more than one speaker. 

Research question. 

## 2. Analysis

### Load Libraries and Themes

```{r libs}
library(tidyverse)
library(tidytuesdayR)
library(tidytext)
library(paletteer)

```

### Load Data

Data found on Github [here.](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-01-14/readme.md) 

```{r}
tuesdata <- tidytuesdayR::tt_load('2025-01-14')

conf2023 <- 
  tuesdata$conf2023 |> 
  mutate(year = "2023")

conf2024 <- 
  tuesdata$conf2024 |> 
  mutate(year = "2024")

```


We will identify the most common words in both years. We will use the `tidytext` [package.](https://cran.r-project.org/web/packages/tidytext/vignettes/tidytext.html) 

### Clean and Wrangle

2023 data
```{r}
conf2023_talks <- 
  conf2023 |>
  select(session_title) |>
  unique()

# tidy table to one word per row
tidy_conf2023_talks <- 
  conf2023_talks |>
  unnest_tokens(output = word, input = session_title)

# remove stop words
tidy_conf2023_talks <-
  tidy_conf2023_talks |>
  anti_join(get_stopwords(), join_by(word))

# group words and count
tidy_conf2023_word_count <- 
  tidy_conf2023_talks |>
  group_by(word) |> 
  summarise(n = n()) |>
  arrange(desc(n)) |> 
  mutate(year = "2023")

```
 
2024 data

```{r}
conf2024_talks <- 
  conf2024 |>
  select(talk_title) |>
  unique()

# tidy table to one word per row
tidy_conf2024_talks <- 
  conf2024_talks |>
  unnest_tokens(output = word, input = talk_title)

# remove stop words
tidy_conf2024_talks <-
  tidy_conf2024_talks |>
  anti_join(get_stopwords(), join_by(word)) 

# group words and count
tidy_conf2024_word_count <- 
  tidy_conf2024_talks |>
  group_by(word) |> 
  summarise(n = n()) |>
  arrange(desc(n)) |> 
  mutate(year = "2024")

tidy_conf_word_count <- 
  rbind(tidy_conf2023_word_count, tidy_conf2024_word_count) 
  
head(tidy_conf_word_count)

```

## 3. Final Plot

```{r}
tidy_conf_word_count |> 
  filter(n >= 5) |> 
  mutate(word = reorder(word, n)) |> 
  ggplot(aes(n, word, fill = year)) + 
  geom_col() + 
  facet_wrap(vars(year), scales = "free_y") +
  scale_fill_paletteer_d("nbapalettes::supersonics_holiday") +
  labs(title = "Most Common Words in posit::conf Talks",
       subtitle = "2023 vs 2024",
       x = "Count", 
       y = "Word") +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "lightgray",
                                colour = "lightgray",
                                size = 0.5, linetype = "solid"), 
        plot.background = element_rect(fill = "lightgray")
  ) 

```


## 4. Conclusion

test