---
title: "Tidy Tuesday, August 18, 2026"
output: html_document
date: "2026-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Libraries
```{r}
library(tidyverse)
library(tidytuesdayR)
library(janitor)

options(scipen = 999)

```

Data
```{r}
tuesdata <- tidytuesdayR::tt_load('2026-08-18')

demo_by_first_language <- tuesdata$demo_by_first_language
demo_by_nationality <- tuesdata$demo_by_nationality
demo_by_reasons <- tuesdata$demo_by_reasons
performance_by_first_language <- tuesdata$performance_by_first_language
performance_by_nationality <- tuesdata$performance_by_nationality

```


## Tidy & Explore

Explore reasons for taking the test. 

Note that data are grouped and the `percent` variable indicates the percentage of population with that reason that received that score. 

Also note that lowest band category is <4, then 4, 4.5, 5, 5.5, and so on through 9, will need to order the categories. 
```{r}
demo_by_reasons 

demo_by_reasons |> 
  group_by(reason) |> 
  summarise(n = n())

demo_by_reasons |> 
  group_by(band) |> 
  summarise(n = n())

demo_by_reasons |> 
  summarise(sum(percent))

demo_by_reasons |> 
  group_by(reason) |> 
  summarise(sum_percent = sum(percent))

demo_by_reasons |> 
  summarise(sum_percent = sum(percent))

demo_by_reasons |> 
  group_by(year) |> 
  summarise(sum_percent = sum(percent))
            
```

- 11 reasons for taking the test (including "Other")
- total percent adds up to 66% (not 100%?)
- maybe make a proportion bar chart and note students that did well, ok, and not good? (i.e. collapse categories)

Change band to factor. 
```{r}
demo_by_reasons <- 
  demo_by_reasons |> 
  mutate(band = factor(band, levels = c('<4', '4', '4.5', '5', '5.5', '6', '6.5', '7', '7.5', '8', '8.5', '9')), 
         percent = round(percent, 3))

```

Does the reason for taking the test affect the score?

faceted distributions
```{r}
demo_by_reasons |> 
  ggplot(aes(x = band, y = percent)) + 
  geom_point() +
  facet_wrap(~reason, ncol = 2)

```
proportion bar chart
```{r}
demo_by_reasons |> 
  ggplot(aes(x = percent, y = reason, fill = fct_rev(band))) + 
  geom_col()
  
```


## Final Visualization

```{r}
final_plot <- 
  demo_by_reasons |> 
  ggplot(aes(x = percent, y = reason, fill = fct_rev(band))) +
  geom_col() + 
  scale_fill_viridis_d(option = "viridis") + 
  labs(x = "Percent in Score Band", 
       y = "", 
       fill = "Score Band", 
       title = "IELTS Exam Scores by Reason for Taking Exam", 
       subtitle = "2022-2025") + 
  theme_minimal()


```


## Save

```{r}
ggsave("test_scores_by_reason.png", 
       plot = final_plot,
       width = 8,       #  inches wide
       height = 5,      # inches high
       units = "in",    # Explicitly use inches
       dpi = 150       # High resolution for print/presentation)
)

```

