---
title: "Tidy Tuesday, August 18, 2026"
output: html_document
date: "2026-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Libraries
```{r}
library(tidyverse)
library(tidytuesdayR)
library(janitor)

options(scipen = 999)

```

Data import.
```{r}
tuesdata <- tidytuesdayR::tt_load('2026-08-18')

demo_by_first_language <- tuesdata$demo_by_first_language
demo_by_nationality <- tuesdata$demo_by_nationality
demo_by_reasons <- tuesdata$demo_by_reasons
performance_by_first_language <- tuesdata$performance_by_first_language
performance_by_nationality <- tuesdata$performance_by_nationality

```


## Tidy & Explore

The International English Language Testing System (IELTS) exam consists of four parts: listening, speaking reading and writing. Each part is scored on a band system from 1 to 9, and the four parts are averaged for a total score. Values for 2022, 2023, and 2024 are included in these datasets. 

There are five datasets, each of which details score breakdown by first language, nationality, and reasons for taking the test. 

Note that in the `demo_by_...` datasets, data are grouped such that the `percent` variable indicates the percentage of population with that reason that received that score. Also note that there are two types of IELTS exams; Academic and General Training and the scores likely differ between them. 

### Performance by First Language

It would make sense that English speakers score the highest, but maybe that only applies to one of the parts (i.e. speaking?)
```{r}
performance_by_first_language |> 
  head()

performance_by_first_language |> 
  group_by(language) |> 
  summarise(n = n()) # 53 languages

performance_by_first_language |> 
  filter(year == '2024-2025' & type == 'Academic') |> 
  ggplot(aes(x = part, y = score, fill = part)) + 
  geom_col() + 
  facet_wrap(~language)

performance_by_first_language |> 
  filter(year == '2024-2025' & type == 'General_Training') |> 
  ggplot(aes(x = part, y = score, fill = part)) + 
  geom_col() + 
  facet_wrap(~language)

```

### Reasons for Taking 

```{r}
demo_by_reasons 

demo_by_reasons |> 
  group_by(reason) |> 
  summarise(n = n())

demo_by_reasons |> 
  group_by(band) |> 
  summarise(n = n())

demo_by_reasons |> 
  summarise(sum(percent))

demo_by_reasons |> 
  group_by(reason) |> 
  summarise(sum_percent = sum(percent))

demo_by_reasons |> 
  summarise(sum_percent = sum(percent))

demo_by_reasons |> 
  group_by(year) |> 
  summarise(sum_percent = sum(percent))
            
```

- lowest band category is <4, then 4, 4.5, 5, 5.5, and so on through 9, will need to order the categories. 
- 11 reasons for taking the test (including "Other")
- total percent adds up to 66% (not 100%?)
- maybe make a proportion bar chart and note students that did well, ok, and not good? (i.e. collapse categories)

Change band to factor. 
```{r}
demo_by_reasons <- 
  demo_by_reasons |> 
  mutate(band = factor(band, levels = c('<4', '4', '4.5', '5', '5.5', '6', '6.5', '7', '7.5', '8', '8.5', '9')), 
         percent = round(percent, 3))

```

Does the reason for taking the test affect the score?

faceted distributions
```{r}
demo_by_reasons |> 
  ggplot(aes(x = band, y = percent)) + 
  geom_point() +
  facet_wrap(~reason, ncol = 1)

```

proportion bar chart
```{r}
demo_by_reasons |> 
  ggplot(aes(x = percent, y = reason, fill = fct_rev(band))) + 
  geom_col()
  

```

Too many categories, re-tool this to be "proportion high-scorers" by reason. 
```{r}
plot_data <- 
  demo_by_reasons |> 
#  filter(type == 'General_Training') |> 
  mutate(high_scorer = if_else(band %in% c('7.5','8','8.5','9'), "high_scorer", "not_high_scorer")) |> 
  group_by(reason, high_scorer) |> # collapse years and test type
  summarise(percent_sum = sum(percent), .groups = "drop_last") |> 
  mutate(proportion = percent_sum/sum(percent_sum))
#  filter(high_scorer == 'high_scorer') |> 
#  arrange(desc(percent_sum))

```


```{r}
plot_data |> 
  ungroup() |> 
  summarise(sum_percent = sum(percent_high))

```


## Final Visualization

```{r}
final_plot <- 
  demo_by_reasons |> 
  ggplot(aes(x = percent, y = reason, fill = fct_rev(band))) +
  geom_col() + 
  scale_fill_viridis_d(option = "viridis") + 
  labs(x = "Percent in Score Band", 
       y = "", 
       fill = "Score Band", 
       title = "IELTS Exam Scores by Reason for Taking Exam", 
       subtitle = "2022-2025") + 
  theme_minimal()

final_plot

```


## Save

```{r}
ggsave("test_scores_by_reason.png", 
       plot = final_plot,
       width = 12,       #  inches wide
       height = 10,      # inches high
       units = "in",    # Explicitly use inches
       dpi = 150       # High resolution for print/presentation)
)

```

