---
title: "Tidy Tuesday, August 18, 2026"
output: html_document
date: "2026-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Libraries
```{r}
library(tidyverse)
library(tidytuesdayR)
library(janitor)

options(scipen = 999)

```

Data
```{r}
tuesdata <- tidytuesdayR::tt_load('2026-08-18')

demo_by_first_language <- tuesdata$demo_by_first_language
demo_by_nationality <- tuesdata$demo_by_nationality
demo_by_reasons <- tuesdata$demo_by_reasons
performance_by_first_language <- tuesdata$performance_by_first_language
performance_by_nationality <- tuesdata$performance_by_nationality

```


## Tidy & Explore

Explore reasons for taking the test. 

Note that data are grouped and the `percent` variable indicates the percentage of population with that reason that received that score. 

Also note that lowest band category is <4, then 4, 4.5, 5, 5.5, and so on through 9, will need to order the categories. 
```{r}
demo_by_reasons 

demo_by_reasons |> 
  group_by(reason) |> 
  summarise(n = n())

demo_by_reasons |> 
  group_by(band) |> 
  summarise(n = n())

demo_by_reasons |> 
  summarise(sum(percent))

demo_by_reasons |> 
  group_by(reason) |> 
  summarise(sum_percent = sum(percent))

demo_by_reasons |> 
  summarise(sum_percent = sum(percent))

demo_by_reasons |> 
  group_by(year) |> 
  summarise(sum_percent = sum(percent))
            
```

- 11 reasons for taking the test (including "Other")
- total percent adds up to 66% (not 100%?)
- maybe make a proportion bar chart and note students that did well, ok, and not good? (i.e. collapse categories)

Does the reason for taking the test affect the score?
```{r}
demo_by_reasons |> 
  mutate(band = factor(band, levels = c('<4', '4', '4.5', '5', '5.5', '6', '6.5', '7', '7.5', '8', '8.5', '9')), 
         percent = round(percent, 3)) |> 
  ggplot(aes(x = reason, y = percent)) + 
  geom_point()


  
```


## Visualize


## Save

```{r}
# ggsave("green_space_change.png", plot = final_plot, 
#   width = 36,       #  inches wide
#   height = 40,      # inches high
#   units = "in",    # Explicitly use inches
#   dpi = 150       # High resolution for print/presentation)
# )

```

