---
title: "Tidy Tuesday, September 22, 2026"
output: html_document
date: "2026-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Libraries
```{r}
library(tidyverse)
library(tidytuesdayR)
library(janitor)

```

Data
```{r}
tuesdata <- tidytuesdayR::tt_load('2026-09-22')
urban <- tuesdata$urban

```

- `greenAreaPerCapitaM2`: green area per capita
- `averageShareOfGreenAreaInCityUrbanAreaPct` Average share of green area in each city in percentage

Want to know cities with highest green area, and which city has gained/lost the most since the first data point in 1990. 

## Tidy & Explore

By region
```{r}
# figure out which cities have values for both 1990 and 2020 (using 2025 limits the data too much)
available_1990_2020 <- 
  urban |> 
  filter(year %in% c(1990, 2020)) |> 
  group_by(cityName) |> 
  summarise(n = n()) |> 
  filter(n == 2)

# calculate difference in area between 2020 and 1990
year_diffs <- 
  urban |> 
  left_join(available_1990_2020, by = "cityName") |> 
  filter(!is.na(n)) |> 
  arrange(cityName, year) |> 
  select(c(sdgRegion, cityName, year, averageShareOfGreenAreaInCityUrbanAreaPct)) |> 
  pivot_wider(names_from = year, values_from = averageShareOfGreenAreaInCityUrbanAreaPct, names_prefix = "year_") |> 
  mutate(diffce_2020_1990 = year_2020 - year_1990) 

# largest increase in green space between 1990 and 2020, by region
largest_increase_green_area <- 
  year_diffs |> 
  select(c(sdgRegion, cityName, diffce_2020_1990)) |> 
  filter(diffce_2020_1990 > 0) |> # only cities where green space increased
  arrange(sdgRegion, desc(diffce_2020_1990)) |> 
  group_by(sdgRegion) |> 
  mutate(row_num = row_number()) |> 
  filter(row_num <= 10) |> 
  ungroup()

# largest decrease
largest_decrease_green_area <- 
  year_diffs |> 
  select(c(sdgRegion, cityName, diffce_2020_1990)) |> 
  filter(diffce_2020_1990 < 0) |> # only cities where green space decreased
  arrange(sdgRegion, desc(diffce_2020_1990)) |> 
  mutate(neg_diffce_2020_1990 = -diffce_2020_1990) |> 
  group_by(sdgRegion) |> 
  mutate(row_num = row_number()) |> 
  filter(row_num <= 10) |> 
  ungroup()

```

Visualize distribution. 
```{r}
year_diffs |> 
  filter(sdgRegion == 'Northern America and Europe') |> 
  ggplot(aes(x = diffce_2020_1990)) + 
  geom_histogram()

```

Largest change (both increase and decrease) 
```{r}
largest_change_green_area <- 
  largest_decrease_green_area |> 
  select(-neg_diffce_2020_1990) |> 
  rbind(largest_increase_green_area) |> 
  mutate(gained_or_lost = if_else(diffce_2020_1990 < 0, "lost", "gained")) |> 
  arrange(sdgRegion, desc(diffce_2020_1990))

```

## Visualize

```{r}
largest_change_green_area |> 
  group_by(sdgRegion) |> 
  ggplot(aes(x = diffce_2020_1990, y = fct_reorder(cityName, diffce_2020_1990), fill = gained_or_lost)) + 
  geom_col(position = "identity") + 
  xlim(-30, 30)+
  # color code manually
  scale_fill_manual(values = c("gained" = "#2ca25f", "lost" = "#8c510a")) + 
  # add labels
  geom_text(aes(label = round(diffce_2020_1990, 0)), 
            vjust = 0.2) + 
  facet_wrap(~sdgRegion, scales = "free_y", ncol = 1) + 
  labs(x = "Pct Green Space Gained or Lost", 
       y = "", 
       fill = "Gained/Lost") + 
  theme_classic()

```

## Conclude

```{r}

```

