---
title: "Tidy Tuesday, September 22, 2026"
output: html_document
date: "2026-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Libraries
```{r}
library(tidyverse)
library(tidytuesdayR)
library(janitor)
library(ggtext)
library(showtext)
library(sysfonts)

```

Fonts
```{r}
font_add_google(
  name = 'Inter', 
  family = 'Inter'
)

showtext_auto()

main_font <- 'Inter'

```

## Data

```{r}
tuesdata <- tidytuesdayR::tt_load('2026-09-22')
urban <- tuesdata$urban

```

- `greenAreaPerCapitaM2`: green area per capita
- `averageShareOfGreenAreaInCityUrbanAreaPct` Average share of green area in each city in percentage

Want to know cities with highest green area, and which city has gained/lost the most since the first data point in 1990. 

## Tidy & Explore

By region
```{r}
# limit to North America and Europe
urban_na_eu <- 
  urban |> 
  filter(sdgRegion == 'Northern America and Europe')

# figure out which cities have values for both 1990 and 2020 (using 2025 limits the data too much)
available_1990_2020 <- 
  urban_na_eu |> 
  filter(year %in% c(1990, 2020)) |> 
  group_by(cityName) |> 
  summarise(n = n()) |> 
  filter(n == 2)

# calculate difference in area between 2020 and 1990
year_diffs <- 
  urban_na_eu |> 
  left_join(available_1990_2020, by = "cityName") |> 
  filter(!is.na(n)) |> 
  arrange(cityName, year) |> 
  select(c(sdgRegion, cityName, year, averageShareOfGreenAreaInCityUrbanAreaPct)) |> 
  pivot_wider(names_from = year, values_from = averageShareOfGreenAreaInCityUrbanAreaPct, names_prefix = "year_") |> 
  mutate(diffce_2020_1990 = year_2020 - year_1990) 

# largest increase in green space between 1990 and 2020, by region
largest_increase_green_area <- 
  year_diffs |> 
  select(c(sdgRegion, cityName, diffce_2020_1990)) |> 
  filter(diffce_2020_1990 > 0) |> # only cities where green space increased
  arrange(sdgRegion, desc(diffce_2020_1990)) |> 
  group_by(sdgRegion) |> 
  mutate(row_num = row_number()) |> 
  filter(row_num <= 10) |> 
  ungroup()

# largest decrease
largest_decrease_green_area <- 
  year_diffs |> 
  select(c(sdgRegion, cityName, diffce_2020_1990)) |> 
  filter(diffce_2020_1990 < 0) |> # only cities where green space decreased
  mutate(neg_diffce_2020_1990 = -diffce_2020_1990) |> 
  arrange(sdgRegion, desc(neg_diffce_2020_1990)) |> 
  group_by(sdgRegion) |> 
  mutate(row_num = row_number()) |> 
  filter(row_num <= 10) |> 
  ungroup()

```

Visualize distribution. 
```{r}
year_diffs |> 
  ggplot(aes(x = diffce_2020_1990)) + 
  geom_histogram()

```

Largest change (both increase and decrease) 
```{r}
largest_change_green_area <- 
  largest_decrease_green_area |> 
  select(-neg_diffce_2020_1990) |> 
  rbind(largest_increase_green_area) |> 
  mutate(gained_or_lost = if_else(diffce_2020_1990 < 0, "lost", "gained")) |> 
  arrange(sdgRegion, desc(diffce_2020_1990))

```

## Plot

```{r}
plot <- 
  largest_change_green_area |> 
  ggplot(aes(x = diffce_2020_1990, y = fct_reorder(cityName, diffce_2020_1990), fill = gained_or_lost)) + 
  geom_col(position = "identity") 

```

## Styling

- color: there are really only two colors - more green space (probably green?) and less green space (probably brown)
- fonts

```{r}
final_plot <- 
  plot + 
  xlim(-30, 30) + # imit axis
  # color code manually
  scale_fill_manual(values = c("gained" = "#3B7A57", "lost" = "#B37D4E")) + 
  # add labels
  geom_text(aes(label = paste0(round(diffce_2020_1990, 0), '%'),
            hjust = if_else(diffce_2020_1990 > 0, -0.2, 1.2))
            ) + 
  labs(x = "Pct Green Space Gained or Lost", 
       y = "", 
       fill = "Gained/Lost", 
       title = "Cities that have <span style='color:#B37D4E;'>Lost</span> or <span style='color:#3B7A57;'>Gained</span> the Most Green Space", 
       subtitle = "North America and Europe, between 1990 and 2020") + 
  theme_classic() + 
  theme(
    text = element_text(family = main_font),
    plot.title = element_markdown(face = "bold", 
                                  size = 18), 
    plot.subtitle = element_markdown(color = "#5E5C5B",   
                                     size = 12,
                                     margin = margin(b = 15)), # Adds a little whitespace cushion below the subtitle
    legend.position = "none", 
    axis.line.x = element_blank()
  )

final_plot

```

## Save

```{r}
ggsave("green_space_change.png", plot = final_plot, 
  width = 7,       #  inches wide
  height = 5,      # inches high
  units = "in",    # Explicitly use inches
  dpi = 150       # High resolution for print/presentation)
)

```

